---
title: "job_search"
format: html
editor: visual
---

## Step 1: Download LinkedIn Connections

LinkedIn \> Settings \> Data Privacy \> Get a Copy of Your Data.

```{r download_li_connections}

library(tidyverse)

connections_raw <- read_csv(
  "drafts/job_search/Connections.csv",
  skip = 2) %>%
  janitor::clean_names()
  
```

## Step 2: Give an LLM all of the companies you have a connection to

```{r}

connections_raw %>%
  count(company, sort = T) %>%
  head(50) %>%
  select(company) %>%
  clipr::write_clip()

connections_raw %>%
  filter(
    !str_detect(company, "Workera"),
    !is.na(company)) %>%
  count(company, sort = T, name = "n_connections") %>%
  slice(601:650) %>% 
  select(company) %>%
  clipr::write_clip()
  write_csv("drafts/job_search/connected_companies.csv")
```

### Prompt

> For every company in this list, go on the web and find their job postings URL. Make sure it is the URL for all open jobs, not a page that only describes what it's like to work there (often called a careers page). Generate a short description of what the company does and save it as a csv with the columns `company_name`, `description`, and `job_postings_url`.
