---
title: "Reddit Sentiment"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(tidytext)
library(progressr)
theme_set(theme_light())
```

```{r read_data}
# Read all CSVs from directory
filenames <- 
  list.files(
    path = "drafts/reddit_sentiment/data", 
    # pattern = "pullpush_run_at_20250420_120233_to_1578586913",
    pattern = "pullpush_", 
    full.names = TRUE) %>%
  set_names()

read_csv(filenames) %>%
  glimpse()
  select(upvote_ratio)

# Read a CSV from a file path. 
# Select a subset of columns and manually specify the types
read_posts <- function(fn) {
  
  p() # Update progress bar 
  
  # Some files do not have upvote_ratio (inexplicably)
  if ("upvote_ratio" %in% names(read.csv(fn, nrows = 0))) {
    read_csv(
      fn,
      col_select = c(
        id, 
        author, 
        created_utc, 
        subreddit_id,
        num_comments, 
        upvotes = score, 
        upvote_ratio, 
        title, 
        selftext, 
        url),
      col_types = cols(
        id = col_character(),
        author = col_character(),
        created_utc = col_number(),
        subreddit_id = col_character(),
        num_comments = col_number(),
        score = col_number(),
        upvote_ratio = col_number(),
        title = col_character(),
        selftext = col_character(),
        url = col_character())) 
  } else {
    read_csv(
      fn,
      col_select = c(
        id, 
        author, 
        created_utc, 
        subreddit_id,
        num_comments, 
        upvotes = score, 
        # upvote_ratio, 
        title, 
        selftext, 
        url),
      col_types = cols(
        id = col_character(),
        author = col_character(),
        created_utc = col_number(),
        subreddit_id = col_character(),
        num_comments = col_number(),
        score = col_number(),
        # upvote_ratio = col_number(),
        title = col_character(),
        selftext = col_character(),
        url = col_character())) %>%
    mutate(upvote_ratio = NA)
  }
}

# Big one: this takes a while  
with_progress({
  p <- progressor(along = 1:length(filenames))
   
  pm_posts_raw <- 
    filenames %>%
    map_df(read_posts, .id = "source_file") %>%
    distinct(id, .keep_all = T) # Remove any duplicate posts
    mutate(date = date(created_utc)) # Add a date col
})

pm_posts_raw %>%
  saveRDS("ProductManagement_all_posts.rds")

pm_posts_raw %>%
  summarise(
    min(date, na.rm = T),
    max(date, na.rm = T))


```

```{r inspect}

pm_posts_raw %>%
  count(date) %>%
  ggplot(aes(date, n)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# TODO: scrape the most recent ones again
pm_posts_raw %>%
  count(date = floor_date(date, "month")) %>%
  ggplot(aes(date, n)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
# For our analysis, let's look only at posts after 2015, because that is when it started taking off

pm_posts_raw %>%
  filter(date >= "2015-01-01") %>%
  ggplot(aes(upvotes, upvote_ratio)) + 
  geom_point() 

pm_posts_raw %>% 
  arrange(desc(upvotes)) %>% 
  View

pm_posts_raw %>%
  filter(date >= "2020-01-01") %>%
  group_by(month = floor_date(date, "month")) %>%
  summarise(
    mean_upvote_ratio = mean(upvote_ratio, na.rm = T), 
    median_upvote_ratio = median(upvote_ratio, na.rm = T)) %>%
  pivot_longer(-month, names_to = "metric", values_to = "value") %>%
  ggplot(aes(month, value, color = metric)) + 
  geom_line()

```

```{r tokenize}
pm_tokens <- 
  pm_posts_raw %>%
  filter(date >= "2015-01-01") %>%
  mutate(text = paste(title, "\\n", selftext)) %>%
  filter(author != "AutoModerator") %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word") 

# confirm uniqueness
pm_posts_raw %>%
  count(id, author, created_utc, num_comments, upvotes, upvote_ratio, url, sort = T)

bing_word_sentiment <- 
  pm_tokens %>%
  inner_join(get_sentiments("bing"), by = "word") 

bing_word_sentiment %>%
  count(date = floor_date(date, "month"), sentiment) %>%
  pivot_wider(names_from = "sentiment", values_from = "n") %>%
  mutate(positive_ratio = positive / (positive + negative)) %>%
  ggplot(aes(date, positive_ratio)) +
  geom_line()

# Now by post
bing_word_sentiment %>%
  count(id, author, date, num_comments, upvotes, upvote_ratio, url, sentiment) %>%
  pivot_wider(names_from = "sentiment", values_from = "n") %>%
  replace_na(list(negative = 0L, positive = 0L)) %>%
  mutate(post_sentiment = if_else(positive > negative, "positive", "negative")) %>%
  count(date = floor_date(date, "month"), post_sentiment) %>%
  group_by(date) %>%
  add_count(wt = n, name = "total") %>%
  mutate(pct = n / total) %>%
  filter(post_sentiment == "positive") %>%
  pivot_longer(cols = c("pct", "total"), names_to = "metric", values_to = "value") %>%
  ggplot(aes(date, value)) + 
  geom_line() + 
  facet_wrap(~metric, ncol = 1, scales = "free_y")


afinn_sentiment <- 
  pm_tokens %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  group_by(
    id, author, created_utc, num_comments, upvotes, upvote_ratio, url) %>%
  summarise(
    net_afinn = sum(value, na.rm = T),
    median_afinn = median(value, na.rm = T),
    mean_afinn = mean(value, na.rm = T)) %>%
  ungroup()


```

## What words come up more now than in the past?

```{r}
afinn_sentiment %>%
  count(date = created_utc) %>%
  ggplot(aes(date, n)) + 
  geom_col()

afinn_sentiment %>%
  count(date = created_utc, positive = net_afinn > 0) %>% 
  group_by(date) %>%
  add_count(name = "total", wt = n) %>%
  mutate(pct_positive = n / total) %>%
  filter(positive) %>%
  ggplot(aes(date, pct_positive)) + 
  geom_col()

afinn_sentiment %>%
  mutate(date = floor_date(created_utc, "week")) %>%
  ggplot(aes(date, net_afinn, group = date)) + 
  geom_boxplot() + 
  scale_y_log10()


afinn_sentiment %>%
  inner_join(
    pm_posts %>% select(id, url),
    by = "id") %>%
  arrange(net_afinn) %>%
  View

```

```{pm_tokens %>%}
  inner_join(get_sentiments("nrc"), by = "word") %>%
  count(sentiment, word) %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>%
  # arrange(sentiment, n)
  ggplot(aes(reorder_within(word, n, sentiment), n)) + 
  geom_col() + 
  facet_wrap(~sentiment, scales = "free_y") + 
  scale_x_reordered() + 
  coord_flip()

pm_tokens %>%
  mutate(date = floor_date(date(created_utc), "week")) %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(date, sentiment) %>%
  pivot_wider(names_from = "sentiment", values_from = n) %>%
  mutate(
    total = positive + negative,
    pct_positive = positive / total) %>%
  ggplot(aes(date, pct_positive)) + 
  geom_line() + 
  geom_smooth() +
  scale_y_continuous(labels = scales::percent) +
  expand_limits(y = c(0, 1))
  

```

1.  **AFINN**: A lexicon that assigns words with a score from -5 (most negative) to 5 (most positive)

2.  **Bing**: Categorizes words as positive or negative

3.  **NRC**: Categorizes words into eight emotions (anger, anticipation, disgust, fear, joy, sadness, surprise, trust) and two sentiments (positive and negative)

4.  **Loughran**: Financial sentiment lexicon particularly useful for analyzing financial texts
